---
title: "Chapter 2, 2.4 Exercises p52-57"
author: "Will Sutton"
date: "18/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

### Conceptual

1. For parts (a)-(d), indicate whether we would generally expect the performance of a flexible statistical learning method to be better or worse than an inflexible method. Justify your answer.

(a) The sample size n is extremely large, and the number of predictors p is small.
(b) The number of predictors p is extremely large, and the number of observations n is small.
(c) The relationship between the predictors and response is highly non-linear.  
(d) The variance of the error terms, i.e. theta^2 = Var(epsilon), is extremely high.

--- 

1. Answers

(a) A flexible method would perform better than an inflexible method because the sample size n is large we have more confidence that our method would be a closer fit of the data.

(b) A flexible method would perform worse than an inflexible method because it would overfit the small number of observations.

(c)

(d)

--- 


2. Explain whether each scenario is a classification or regression problem, and indicate whether we are most interested in inference or prediction. Finally, provide n and p. 

(a) We collect a set of data on the top 500 firms in the US. 

For each firm we record:
- profit
- number of employees
- industry
- the CEO salary

We are interested in understanding which factors affect CEO salary. 

(b) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. 

For each product we have recorded:
- whether it was a success or failure
- price charged for the product
- marketing budget
- competition price
- ten other variables.  

(c) We are interested in predicting the % change in the USD/Euro exchange rates in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. 

For each week we record:
- the % change in the USD/Euro
- the % change in the US market
- the % change in the British market 
- the % change in the German market

---

2. Answers

(a) regression
inference
n = 500
p = 3 

(b) classification
prediction
n = 20
p = 13

(c) regression
prediction
n = 52
p = 3

---

3. We now revisit the bias-variance decomposition.

(a) Provide a sketch of typical:
- (squared) bias
- variance
- training error
- test error
- Bayes (or irreducible) error curves
on a single plot, as we go from less flexible statistical learning methods towards more flexible approaches. 

The x-axis should represent the amount of flexibility in the method, and the y-axis should represent the values for each curve. there should be five curves. Make sure to label each one. 

(b) Explain why each of the five curves has the shape displayed in part (a). 

---

3. Answers

(a)

(b)
---


4. You will now think of some real-life applications for statistical learning.

(a) Describe three real-life applications in which classification might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer. 

(b) Describe three real-life applications in which regression might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction? Explain your answer. 

(c) Describe three real-life applications in which cluster analysis might be useful.

---

4. Answers

(a) 
Determining if a cell is cancerous or not, prediction
Determining if an email is spam or not, prediction
Understanding what factors make a successful product, inference

(b)
Predicting the movement in a stock price, prediction
Predicting the weight of a person, prediction
Understanding features that make a car more fuel efficient, inference 


(c)
Describing and segmenting groups of consumers
Understanding different strains of a virus
Movie recommendations

---


5. What are the advantages and disadvantages of a very flexible (versus a less flexible) approach for regression or classification? 
Under what circumstances might a more flexible approach be preferred to be a less flexible approach? 
When might a less flexible approach be preferred?

---

5. Answers

The advantages of a flexible approach is a better fit to the data over a less flexible approach.

A less flexible approach may be preferred when the number of observations is small or when there is a danger of over fitting the data.

---

6. Describe the differences between a parametric and a non-parametric statistical learning approach. What are the advantages of a parametric approach to  regression or classification (as opposed to a non-parametric approach)? What are its disadvantages?

---

6. Answers

---


7. The table below provides a training data set containing six observations, three predictors, and one qualitative response variable.

Obs. | X1 | X2 | X3 | Y
------------------------
1 | 0 | 3 | 0 | Red
2 | 2 | 0 | 0 | Red
3 | 0 | 1 | 3 | Red
4 | 0 | 1 | 2 | Green
5 | -1 | 0 | 1 | Green
6 | 1 | 1 | 1 | Red

Suppose we wish to use this data set to make a prediction for Y when X1 = X2 = X3 = 0 using K-nearest neighbours. 

(a) Compute the Euclidean distance between each observation and the test point, X1 = X2 = X3 = 0.
(b) What is our prediction with K = 1? why?
(c) What is our prediction with K = 3? why?
(d) If the Bayes decision boundary in this problem is highly non-linear, then would we expect the best value for K to be large or small? Why?

---

7. Answers

(a)
3
2
sqrt(10)
sqrt(5)
sqrt(2)
sqrt(3)

(b)

(c)

(d)
---



### Applied

8. the exercise relates to the `College` data set, which can be found in the file `College.csv`. It contains a number of variables for 777 different universities and colleges in the US. Variable descriptions available in the book.

(a) Load the data into R using `read.csv()`
(b) Look at the data using the `fix()` function
(c) 
i) Use the `summary()` function to produce a numerical summary
ii) Use the `pairs()` function to produce a scatter plot matrix
iii) Use the `plot()` function to produce side-by-side box plots of `Outstate` versus `Private`
iv) Create a new variable called `Elite` by binning the `Top10perc`
v) Use the `hist()` function to produce some histograms with differing numbers of bins for a few of the quantitative variables.
vi) Continue to exploring the data and provide a brief summary of what you discover.


```{r}
# (a)
data <- read.csv("data/College.csv")

# (b)
fix(data)

# (c)
# i
summary(data)

# ii
pairs(data)

# iii

```


9. This exercise involves the `Auto` data set, make sure that the missing values have been removed from the data.

(a) Which of the predictors are quantitative, and which are qualitative?
(b) What is the range of each quantitative predictor? Use the `range()` function.
(c) What is the mean and standard deviation of each quantitative predictor?
(d) Now remove the 10th through 85th observations. What is the range, mean, and the standard deviation of each predictor in the subset of the data that remains?
(e) Using the full data set, investigate the predictors graphically, using scatter plots or other tools of your choice. Create plots highlighting the relationships among the predictors. Comment on your findings.
(f) suppose that we wish to predict gas mileage(`mpg`) on the basis of the other variable. Do your plots suggest that any other variables might be useful in predicting `mpg`? Justify your answer.


10. This exercise involves the `Boston` data set.

(a) Load the data from the `Mass` library, read about the data set using `?Boston`, how many columns? how many columns? what do the rows and columns represent?
(b) Make some pairwise scatter plots of the predictors (columns) in this data set. Describe your findings.
(c) Are any of the predictors associated with per capita crime rate? If so, explain the relationship.
(d) Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teach ratios? Comment on the range of each predictor.
(e) How many of the suburbs in the data set bound the Charles river?
(f) What is the median pupil-teach ratio among the towns in this data set?
(g) Which suburb of Boston has lowest median value of owner-occupied homes? what are the values of the other predictors for that suburb, and how do those value compare to the overall ranges for those predictors? Comment on your findings.
(h) In this data set, how many of the suburbs average more than seven rooms per dwelling? More than eight rooms per dwelling? Comment on the suburbs that average more than eight rooms per dwelling.

```{r setup, include=FALSE}
#install.packages("MASS")
library(MASS)
```
